Case for Support

### 2021-11-17 application/backgroundToTheResource.tex
Updates from Overleaf

To understand the neural basis of behaviour, experiments need to control and quantify animal behaviour while recording, and often manipulating, neural activity. 
%
Powerful experimental designs require behavioural feedback (e.g.\ reward, lights, sounds, actuated
devices) or neural perturbation (e.g.\ using light-sensitive polarising
molecules) that depend on real-time behaviour or neural state.
%
It has also become increasingly apparent that the neural signals that
drive complex behaviour are manifest in a broad population of neurons rather
than in single cells. Current neural recording emphasises high-dimensional
signals from 100s to 1000s of neurons monitored using multi-electrode arrays or cellular-resolution imaging, with a wide array of probe configurations and neural targets.
%
\textbf{This complexity of experiments limits the potential for database-driven progress in systems neuroscience}. The number of possible experiments is astronomical, and new questions are typically answered by designing and conducting new experiments rather than mining data from old ones.

At present, progress in the field is held back by the challenge of designing and implementing new experimental protocols, and the difficulty of integrating state-of-the-art data processing and neuroinformatics into custom experimental designs.
%
\textbf{Reproducibility}, in particular, depends on the availability of
standardised tools for experimental control and analysis, shared between
laboratories \citep{baker16,ioannidis05}. Although such tools exist in
genetics, astronomy, physics and medicine 
\citep{fishEtAl16,abdallaEtAl18,cern18,dickinsonEtAl16,bycroftEtAl18}, they are
mostly lacking in neuroscience, with documented consequences for the
reliability of ``discoveries'' \citep{baker16,botvinikNezerEtAl20,buttonEtAl13}.

@@

The proposed project will \textbf{enhance, reinforce and support} this key neuroinformatics
resource to streamline \textbf{reproducible experimental design, data collection and analysis}, while making \textbf{standardised implementations of cutting-edge machine-learning methodologies} widely available to the community, potentially creating the most powerful experimental
control and analysis tool available to neuroscientists, psychologists and
ethologists worldwide.

>>>>

To understand the neural basis of behaviour, experiments need to control and quantify animal behaviour while recording, and often manipulating, neural activity. 
%
Powerful experimental designs require behavioural feedback (e.g.\ rewards, sounds, actuated
devices) or neural perturbation (e.g.\ using light-sensitive polarising
molecules) that depend on real-time behaviour or neural state.
%
It has also become increasingly clear that the neural signals that
drive complex behaviour are manifest in a broad population of neurons rather
than in single cells. Neural recording now emphasises high-dimensional
signals from 100s to 1000s of neurons monitored using multi-electrode arrays or cellular-resolution imaging, with a wide array of probe configurations and neural targets.
%
\textbf{This complexity of experiments limits the potential for database-driven progress in systems neuroscience}. The number of possible experiments is astronomical, and new questions are typically answered by designing new experiments rather than mining data from old ones.

Progress is thus held back by the challenges of designing and implementing new experimental protocols, and of integrating state-of-the-art data processing and neuroinformatics into custom experimental designs.
%
\textbf{Reproducibility}, in particular, requires standardised tools for experimental control and analysis, shared between
laboratories \citep{baker16,ioannidis05}. Although such tools exist in genetics, astronomy, physics and medicine (Fish et al., 2016; Abdalla et al., 2018, CERN Education, Communications and Outreach Group, 2018, Dickinson et al., 2016, Bycroft et al., 2018), they are scarce in neuroscience, with documented consequences for the reliability of ``discoveries'' (Baker, 2016; Botvinik-Nezer et al., 2020; Button et al., 2013).

@@

The proposed project will \textbf{enhance, reinforce and support} this key neuroinformatics
resource to streamline \textbf{reproducible experimental design, data collection and analysis}, while making \textbf{standardised implementations of cutting-edge machine-learning methodologies} widely available to the community. This will potentially create the most powerful experimental
control and analysis tool available to neuroscientists, psychologists and
ethologists worldwide.

### 2021-11-17 application/communityDemand.tex
Updates from Overleaf

The ongoing AI revolution makes integration of these
technologies into Bonsai extremely timely. The proposed software infrastructure will facilitate the integration of novel ML methods into the Bonsai
ecosystem. This integration will unlock new value
and scientific directions from the rich behavioral and neural
experimental data that Bonsai already allows scientists to collect. This
integration will also provide experimentalists in the biological
sciences who lack programming skills the opportunity to combine real-time ML with
experimental control in neuroscience. To our knowledge
this has not been done before on such a comprehensive scale, and will, we believe, make Bonsai into a fundamental and transformative technology for
the next generation of animal behavioural research.

>>>>

The ongoing AI revolution makes integration of these
technologies into Bonsai extremely timely. The proposed software infrastructure will facilitate the integration of novel ML methods into the Bonsai
ecosystem. This integration will unlock new value
and scientific directions from the rich behavioural and neural
experimental data that Bonsai already allows scientists to collect. This
integration will also provide experimentalists in the biological
sciences who lack programming skills the opportunity to combine real-time ML with
experimental control in neuroscience. To our knowledge
this has not been done before on such a comprehensive scale, and will, we believe, make Bonsai into a fundamental and transformative technology for
the next generation of animal behavioural research.

### 2021-11-17 application/otherResources.tex
Updates from Overleaf

The space of technologies serving experimental control and behaviour monitoring
is large, and is traditionally occupied either by domain-specific graphical
user interfaces for control and recording of specific devices and experiment
types (e.g.\ Open Ephys
GUI\footnote{https://open-ephys.org/gui/},
Miniscope DAQ
Software\footnote{https://github.com/Aharoni-Lab/Miniscope-DAQ-QT-Software})
or by real-time control frameworks for specifying task logic using state
machine or similar formalisms (e.g. NIMH ML\footnote{https://monkeylogic.nimh.nih.gov/},
pyControl\footnote{https://pycontrol.readthedocs.io/en/latest/},
Autopilot\footnote{https://docs.auto-pi-lot.com/en/latest/},
Sanworks\footnote{https://sanworks.io/index.php}).
These dedicated interfaces are typically very comfortable for experimenters
operating within the specific domain that the tool is designed for, but tend to
become unwieldy when there is a need to introduce a new device or variation of
a task which is outside the scope of the framework. The alternative is to use a
more general programming language like Python or MATLAB, with the disadvantage
of the code being harder to understand, maintain, and change. Programming
languages like LabVIEW straddle the middle ground and provide a high-level and
flexible visual interface for composing data acquisition and control systems
themselves. Unlike Bonsai, however, the graphical elements of LabVIEW are
heterogeneous and very fine grained, creating the need for long and complicated
logical structures to implement even a simple experimental control system. By
providing an extremely simple, yet flexible, visual syntax, Bonsai opens the
opportunity even for complete non-programmers to design and successfully
customize relatively complex experiments from the ground up. It is mostly this
capability which has made Bonsai so attractive as a standard tool in
experimental neuroscience.

>>>>

The large field of technologies serving experimental control and behaviour monitoring is traditionally occupied either by domain-specific graphical
user interfaces for control and recording of specific devices and experiment
types (e.g.\ Open Ephys
GUI\footnote{https://open-ephys.org/gui/},
Miniscope DAQ
Software\footnote{https://github.com/Aharoni-Lab/Miniscope-DAQ-QT-Software})
or by real-time control frameworks for specifying task logic using state
machine or similar formalisms (e.g. NIMH ML\footnote{https://monkeylogic.nimh.nih.gov/},
pyControl\footnote{https://pycontrol.readthedocs.io/en/latest/},
Autopilot\footnote{https://docs.auto-pi-lot.com/en/latest/},
Sanworks\footnote{https://sanworks.io/index.php}).
These dedicated interfaces are typically very comfortable for experimenters in the specific domain for which the tool is designed, but can become unwieldy with the introduction of a new device or task variation from outside their usual scope. Alternatively, one can use a more general programming language such as Python or MATLAB, with the disadvantage
of the code being harder to understand, maintain, and change. Programming
languages like LabVIEW straddle the middle ground and provide a high-level, flexible visual interface for composing data acquisition and control systems. Unlike Bonsai, however, the graphical elements of LabVIEW are
heterogeneous and very fine grained, thus requiring long and complex
logical structures to implement even a simple experimental control system. By providing an extremely simple, yet flexible visual syntax, Bonsai provides the
opportunity for even complete non-programmers to design and successfully
customise relatively complex experiments from the ground up. It is this
capability in particular which has made Bonsai such an attractive standard tool in
experimental neuroscience.

### 2021-11-17 application/userEngagement.tex
Updates from Overleaf

The proposed development will facilitate interfacing with other programming languages
such as Python, R, and MATLAB. This will therefore attract a whole new
community of Bonsai users with expertise in software development. To fully integrate this new community and provide them with the full power of Bonsai,
we plan to organize hackathons to accelerate integration of their existing data
analysis and ML algorithms and methods into Bonsai using the new
machine learning package.

>>>>

The proposed development will facilitate interfacing with other programming languages
such as Python, R, and MATLAB. This will therefore attract a whole new
community of Bonsai users with expertise in software development. To fully integrate this new community and provide them with the full power of Bonsai,
we plan to organise hackathons to accelerate integration of their existing data
analysis and ML algorithms and methods into Bonsai using the new
machine learning package.
